---
title: "STT2860: Final Presentation"
author: "Author: Lauren Rusnak"
date: "`r format(Sys.time(), '%B %d, %Y @ %I:%M %p')`"
output: ioslides_presentation
---

```{r global_options, include = FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  comment = NA
)
```

## 

### **Packages Used**

I used tidyverse for data manipulations.
```{r, echo = TRUE}
library(tidyverse)
```

I used mapdata to see if global trends existed in the data.
```{r, echo = TRUE}
library(mapdata)
```


## 

### **Environment Data - UNdata**

I found data from the UNdata website (https://data.un.org/).
The dataset was last updated November 2021.

**Variables in the Dataset:**

* `CO2 emission estimates`: country for which the CO2 emission is being estimated
* `Year`: year of the CO2 estimate
* `Series`: Total CO2 Emissions or CO2 Emissions per capita
* `Value`: The number in thousand metric tons of CO2 or the number in metric tons of CO2


```{r}
CO2_data <- read_csv("Carbon Dioxide Emission Estimates Raw Data.csv")
```

## 

### **Data Manipulations**

* `pivot_wider()`: Manipulated the data so that the total emissions and emission per capita were in different columns   
* `rename()`: Renamed the columns to better signify data and into more R friendly format
* `filter()`: Filtered the data so I only had data from the most recent year: 2017
* `transmute()` & `case_when()`: Changed the names of countries so that the would match the names in mapdata  
  
  
```{r}
CO2_data_clean <-
  pivot_wider(data = CO2_data,
              names_from = "Series",
              values_from = "Value") %>%
  rename("Country" = "CO2 emission estimates", "Total_Emissions" = "Emissions (thousand metric tons of carbon dioxide)", "Emissions_per_capita" = "Emissions per capita (metric tons of carbon dioxide)") %>%
  filter(Year == 2017) %>%
   
  transmute(region = case_when(
    Country == "United States of America" ~ "USA",
    Country == "Bolivia (Plurin. State of)" ~ "Bolivia",
    Country == "Venezuela (Boliv. Rep. of)" ~ "Venezuela",
    Country == "United Kingdom" ~ "UK",
    Country == "Czechia" ~ "Czech Republic",
    Country == "Russian Federation" ~ "Russia",
    Country == "Iran (Islamic Republic of)" ~ "Iran",
    Country == "Viet Nam" ~ "Vietnam",
    Country == "Syrian Arab Republic" ~ "Syria",
    Country == "Côte d’Ivoire" ~ "Ivory Coast",
    Country == "Dem. Rep. of the Congo" ~ "Democratic Republic of the Congo",
    Country == "Dem. People's Rep. Korea" ~ "North Korea",
    Country == "Brunei Darussalam" ~ "Brunei",
    TRUE ~ Country
    ),
    Year = Year,
    Total_Emissions = Total_Emissions,
    Emissions_per_capita = Emissions_per_capita)

```


## 
### **CO2 Total Emissions**

```{r, fig.align = "center"}
CO2_data_clean %>%
  ggplot(aes(x = Total_Emissions)) +
  geom_boxplot() +
  theme(aspect.ratio = 5/15,
        plot.title = element_text(hjust = 0.5),
        plot.caption = element_text(hjust = 0)) +
  labs(title = "Total CO2 Emissions of Countries in 2017",
       x = "Total CO2 Emissions (thousand metric tons of CO2)",
       caption = "
There are several outliers that contribute greatly to the total CO2 emissions. 
The biggest offenders: China, USA, India, Russia, and Japan")

```

##

```{r}
mapdata <- map_data("world")

mapdata_CO2 <- left_join(mapdata, CO2_data_clean, "region")


```


```{r, fig.align = "center"}
ggplot(mapdata_CO2) +
  geom_polygon(aes(x = long, y = lat, group = group, fill = Total_Emissions)) +   
  scale_fill_gradient2(midpoint = 5000000, low = "lightyellow", mid = "orange", high = "red2") +
  labs(title = "CO2 Total Emissions of each Country",
       fill = "Total CO2 Emissions 
(Thousand Metric Tons)",
caption = "
Countries with the largest populations (China, USA, India, etc.) appear to have the highest total 
carbon dioxide emissions."
       ) +
  theme(legend.position = "right",
        plot.title = element_text(hjust = 0.5),
        plot.caption = element_text(hjust = 0),
        aspect.ratio = 7/14)
  
```

```{r, results = "hide"}
print(as_tibble(arrange(CO2_data_clean, desc(Total_Emissions))), n = 5)
```

What would happen if their populations were taken into account?

## 
### **CO2 Emissions per capita**

```{r}
CO2_data_clean %>%
  ggplot(aes(x = Emissions_per_capita)) +
  geom_boxplot() +
  theme(aspect.ratio = 5/15,
        plot.title = element_text(hjust = 0.5),
        plot.caption = element_text(hjust = 0)) +
  labs(title = "CO2 Emissions per capita in 2017",
       x = "CO2 Emissions per capita (metric tons of CO2)",
       caption = "
There are several outliers, but most countries do not overlap. The only countries that appear in the top eleven countries
for total emission and emissions per capita lists are Canada, Saudi Arabia, and the US. The biggest offenders with 
regard to carbon dioxide emission per capita are: Qatar, Curaçao, Kuwait, United Arab Emirates, Gibraltar, and Bahrain.")
```
```{r, results = "hide"}
print(as_tibble(arrange(CO2_data_clean, desc(Emissions_per_capita))), n = 10)
```


##

```{r, fig.align = "center"}
ggplot(mapdata_CO2) +
  geom_polygon(aes(x = long, y = lat, group = group, fill = Emissions_per_capita)) +   
  scale_fill_gradient2(midpoint = 15, low = "light yellow", mid = "orange", high = "red2") +
  labs(title = "CO2 Emissions per capita of each Country",
       fill = "Emissions per capita 
(Metric Tons)",
caption = "
The regions that have the highest CO2 emissions per capita appear to be The Middle East, 
especially the Arabian Penisula, as well as North America."
       ) +
  theme(legend.position = "right",
        plot.title = element_text(hjust = 0.5),
        aspect.ratio = 8/15)
  
```

##
### **Other Lines of Inquiry**

* How has carbon dioxide emissions changed overtime?
* Is there correlation between 1st/2nd/3rd world countries and CO2 emissions?
* How has carbon dioxide emissions changed since COVID?